### 2023-06-23 emnlp2023-latex/fig/docsplit.tex
changes

\begin{tikz}
\small
\node{
Deep learning is part of a broader family 
};
\end{tikz}