### 2023-06-23 emnlp2023-latex/fig/docsplit.tex
changes

\begin{tikzpicture}
\small
    \node[draw,thick,rounded corners=.05in] (doc) [label=above:original training document ($x$)]{
\tiny
    \parbox{2.5in}{
    \color{blue}
    Deep learning is part of a broader family  of machine learning methods based on artificial neural networks with representation learning.
    \color{magenta}
    Artificial neural networks were inspired by information processing and distributed communication nodes in biological systems.
    \color{blue}
    Artificial neural networks have various differences from biological brains.
    \color{magenta}
    Specifically, artifical neural networks tend to be static and symbolic, whereas the biological brain of most living organisms is dynamic and analog.
    }
};
\node[below=of doc,node distance=0.0in] (placeholder) {};
    \node[draw,thick,rounded corners=.05in,left of=placeholder,node distance=0.1in,anchor=north east]  (x1) [label=above:document summary ($x_1^+$)] {
    \tiny
    \color{blue}
    \parbox{1.2in}{
    Deep learning is part of a broader family  of machine learning methods based on artificial neural networks with representation learning.
    Artificial neural networks have various differences from biological brains.
}};
    \node[draw,thick,rounded corners=.05in,right of=placeholder,node distance=0.1in, anchor=north west] (x2) [label=above:document summary ($x_2^+$)]{
    \tiny
    \color{magenta}
    \parbox{1.2in}{
    Artificial neural networks were inspired by information processing and distributed communication nodes in biological systems.
    Specifically, artifical neural networks tend to be static and symbolic, whereas the biological brain of most living organisms is dynamic and analog.
}
};

    \node[above of=x1, node distance=0.51in] (x1a) {};
    \node[above of=x2, node distance=0.61in] (x2a) {};
    \draw[->,thick] (doc) -- (x1a);
    \draw[->,thick] (doc) -- (x2a);
\end{tikzpicture}